---
title: "Clean Survey Data"
format: 
  html:
    page-layout: full
    self-contained: true
    toc: true
    toc-depth: 3
execute: 
  warning: false
editor_options: 
  chunk_output_type: console
---

## Code Status

Complete - pending review

## Purpose

Prepare survey data for analysis regarding research Q's

1. Create dependent categorical variables:
- Largest apprehension between group discussions and solo conversations
- Most frequently used transportation between rideshare and public transit

2. Basic cleaning of independent open-text variables for initial analysis:
- Remove punctuation and capitalization

## Setup

```{r}
library(tidyverse)

# Point to data folders in the repo
path_org <- 'data_org'
path_clean <- 'data_clean'

# Load in data files

data_qualtrics <- read_csv(here::here(
  path_org, 'PRCAQualtricsExport_FileC.csv'))

data_prolific <- read_csv(here::here(
  path_org, c('PRCAProlificExport_FileA.csv', 'PRCAProlificExport_FileB.csv')))

# Join by prolific ID number
survey <- left_join(data_qualtrics, data_prolific, by = c("Q0" = "Participant id"))

# Rename columns
survey <- survey |> 
  rename(
    q1_dislike_groupdiscussion = Q1,
    q2_comfort_groupdiscussion = Q2,
    q3_nervous_groupdiscussion = Q3,
    q4_like_groupdiscussion = Q4,
    q5_tense_newppl_groupdiscussion = Q5,
    q6_relaxed_groupdiscussion = Q6,
    q13_nervous_newppl_convo = Q13,
    q14_nofear_speaking_convo = Q14,
    q15_nervous_convo = Q15,
    q16_relaxed_convo = Q16,
    q17_relaxed_newppl_convo = Q17,
    q18_afraid_convo = Q18,
    q26_public_transit_3months = Q26,
    q27_public_transit_dayuse = Q27,
    q28_rideshare_3months = Q28,
    q29_rideshare_dayuse = Q29,
    q20_license = Q20,
    q21_car_access = Q21,
    q_advice_text = Q18.1,
    q_transport_pref_text = Q19,
    dem_age = Age,
    dem_sex = Sex,
    dem_country = 'Country of residence',
    dem_student = 'Student status',
    dem_employment = 'Employment status')
```

## Missing data

For each research Q, we will be restricted to participants that answered the
corresponding open-text question.

```{r}
survey |> summarise(
  n_apprehension = sum(!is.na(q_advice_text)),
  n_transportation = sum(!is.na(q_transport_pref_text))
  )
```

This restricts our initial sample size to n = 165 for both research questions.

```{r}
# Missing data for those who provided q_advice_text
survey |>
  filter(!is.na(q_advice_text)) |>
  summarise(across(everything(), \(x) sum(is.na(x)))) |>
  pivot_longer(everything(), names_to = "col", values_to = "n_na") |>
  filter(n_na > 0) |>
  arrange(desc(n_na))

# Missing data for those who provided q_transport_pref_text
survey |>
  filter(!is.na(q_transport_pref_text)) |>
  summarise(across(everything(), \(x) sum(is.na(x)))) |>
  pivot_longer(everything(), names_to = "col", values_to = "n_na") |>
  filter(n_na > 0) |>
  arrange(desc(n_na))

# Missing PRCA data within the apprehension research sample
survey |>
  filter(!is.na(q_advice_text)) |>
  rowwise() |>
  mutate(n_na = sum(is.na(c_across(q1_dislike_groupdiscussion:q18_afraid_convo)))) |>
  ungroup() |>
  count(n_na) |>
  arrange(desc(n_na))
```

q_29_rideshare_dayuse is critical for engineering our outcome variable for
the transportation methods research question. The 7 NA observations will need
to be excluded, however those participants can be kept for the apprehension
research question.

There are also 4 observations in the apprehension research question sample that
are missing a value from 1 PRCA question each. While this is a minimal amount
of missing data, it is predominantly missing from solo conversation Q's, and may
bias our engineered outcome of 'biggest_fear' as a result. These 4 observations
will be removed from the apprehension sample.

Removals for both samples will occur after all processing steps before data is
exported.

## Processing for apprehension research Q

```{r}
# Identify forward and reverse-coded Q's (Higher scores = more apprehensive)
forward_code <- c(
  "q1_dislike_groupdiscussion",
  "q3_nervous_groupdiscussion",
  "q5_tense_newppl_groupdiscussion",
  "q13_nervous_newppl_convo",
  "q15_nervous_convo",
  "q18_afraid_convo"
  )
reverse_code <- c(
  "q2_comfort_groupdiscussion",
  "q4_like_groupdiscussion",
  "q6_relaxed_groupdiscussion",
  "q14_nofear_speaking_convo",
  "q16_relaxed_convo",
  "q17_relaxed_newppl_convo"
  )
# Set numeric values for Likert scales
likert_forward <- c(
  "Strongly disagree"          = 1,
  "Somewhat disagree"          = 2,
  "Neither agree nor disagree" = 3,
  "Somewhat agree"             = 4,
  "Strongly agree"             = 5
)
likert_reverse <- c(
  "Strongly disagree"          = 5,
  "Somewhat disagree"          = 4,
  "Neither agree nor disagree" = 3,
  "Somewhat agree"             = 2,
  "Strongly agree"             = 1
)
# Convert to numeric
survey <- survey |>
  mutate(
    across(all_of(forward_code), ~ likert_forward[.x]),
    across(all_of(reverse_code), ~ likert_reverse[.x])
  )
# Identify group discussion vs. solo conversation Q's
group_discussion <- c(
  "q1_dislike_groupdiscussion",
  "q2_comfort_groupdiscussion",
  "q3_nervous_groupdiscussion",
  "q4_like_groupdiscussion",
  "q6_relaxed_groupdiscussion"
  )
solo_conversation <- c(
  "q13_nervous_newppl_convo",
  "q14_nofear_speaking_convo",
  "q15_nervous_convo",
  "q16_relaxed_convo",
  "q17_relaxed_newppl_convo",
  "q18_afraid_convo"
  )
# Calculate biggest fear
survey <- survey |> 
  mutate(
    
#   Sum of group apprehension Q's
    group_apprehension = rowSums(
      pick(all_of(group_discussion)), na.rm = TRUE),
    
#   Sum of solo apprehension Q's
    solo_apprehension = rowSums(
      pick(all_of(solo_conversation)), na.rm = TRUE),

#   Assign greatest of the two values (equal becomes NA)
    biggest_fear = case_when(
      group_apprehension > solo_apprehension ~ "group_apprehension",
      solo_apprehension > group_apprehension ~ "solo_apprehension",
      .default = NA)
  )
```

## Processing for transportation research Q

```{r}
# Set numeric values for transportation variables
usage_monthly <- c(
  "Never"                     = 0,
  "0-1 days a month"          = 1,
  "2-4 days a month"          = 3,
  "4-8 days a month"          = 6,
  "8 or more days a month"    = 8
  )
usage_daily <- c(
  "1-2 rides in a typical day"       = 1.5,
  "3-4 rides in a typical day"       = 3.5,
  "5-6 rides in a typical day"       = 5.5,
  "7 or more rides in a typical day" = 7
  )

# Calculate highest usage transport method
survey <- survey |>
  mutate(
    
#   Convert 3-month usage variables to numeric
    across(c(q26_public_transit_3months, q28_rideshare_3months), ~ usage_monthly[.x]),
    
#   Convert typical daily usage variables to numeric
    across(c(q27_public_transit_dayuse, q29_rideshare_dayuse), ~ usage_daily[.x]),

#   Multiply usage over past 3 months by typical daily usage (usage count estimate)
    public_transit_count = q26_public_transit_3months * q27_public_transit_dayuse,
    rideshare_count = q28_rideshare_3months * q29_rideshare_dayuse,

#   Assign highest usage transport method (equal if BA)
    highest_transport = case_when(
      public_transit_count > rideshare_count ~ "public_transit",
      rideshare_count > public_transit_count ~ "rideshare",
      .default = NA)
    )
```

## open text cleaning

This script will perform only basic cleaning of the text by removing punctuation and capitalization. Further cleaning approaches such as removing stop words or stemming/lemmatization will be considered in other scripts.

```{r}
survey <- survey |> mutate(
    q_advice_text =
      str_to_lower(str_remove_all(q_advice_text, "[[:punct:]]")),
  q_transport_pref_text =
    str_to_lower(str_remove_all(q_transport_pref_text, "[[:punct:]]"))
  )
```

## Filter and export

```{r}
# Full cleaned data for EDA
survey |> write_csv(here::here(path_clean, "survey_clean.csv"))

survey |> filter(
  
# Remove observations with NA values
  !is.na(q_advice_text),
  if_all(q1_dislike_groupdiscussion:q18_afraid_convo, \(x) !is.na(x))) |>

# Remove redundant columns and save
  select(q_advice_text, biggest_fear) |>
  write_csv(here::here(path_clean, "survey_apprehension.csv"))

# Data for transport research Q
survey |> filter(
  
# Remove observations with NA values
  !is.na(q_transport_pref_text),
  if_all(q26_public_transit_3months:q29_rideshare_dayuse, \(x) !is.na(x))) |>
  
# Remove redundant columns and save
  select(q_transport_pref_text, highest_transport) |>
  write_csv(here::here(path_clean, "survey_transportation.csv"))
```
