---
title: "EDA of dependent variables"
format: 
  html:
    page-layout: full
    self-contained: true
    toc: true
    toc-depth: 3
execute: 
  warning: false
editor_options: 
  chunk_output_type: console
---

## Code Status

In Progress (Expected to need updating upon receiving the full dataset)

## Purpose

Conduct EDA on both dependent variables (greatest apprehension & most frequently used transportation method)

- Prepare tables/figures for memo on the distribution of dependent variables

- Elaborate on considerations for modelling informed by these findings

## Setup

```{r}
library(tidyverse)
path_clean <- 'data_clean'
survey <- read_csv(here::here(path_clean, "survey_clean.csv"),
                   show_col_types = FALSE)

# Apprehension sample
survey_ap <- survey |> filter(
  !is.na(q_advice_text),
  if_all(q1_dislike_groupdiscussion:q18_afraid_convo, \(x) !is.na(x)))

# Transportation sample
survey_ts <- survey |> filter(
  !is.na(q_transport_pref_text),
  if_all(q26_public_transit_3months:q29_rideshare_dayuse, \(x) !is.na(x)))
```

## EDA of apprehension scores

*Shape of both distributions*

```{r}
survey_ap |>
  pivot_longer(
    cols = c(solo_apprehension, group_apprehension),
    names_to = "type",
    values_to = "score"
  ) |>
  ggplot(aes(x = score, fill = type)) +
  geom_density(alpha = 0.5)
```

```{r}
r2 <- summary(lm(group_apprehension ~ solo_apprehension, data = survey_ap))$r.squared

ggplot(survey_ap, aes(x = solo_apprehension, y = group_apprehension)) +
  geom_point() +
  geom_smooth(method = "lm") +
  annotate("text", x = Inf, y = Inf, hjust = 1.1, vjust = 1.5,
           label = paste0("R\u00b2 = ", round(r2, 2)))
```

```{R}
survey_ap |>
  mutate(biggest_fear = biggest_fear |>
           replace_na("NA") |>
           factor(levels = c("solo_apprehension", "NA", "group_apprehension"))) |>
  ggplot(aes(x = biggest_fear)) +
  geom_bar()
```

## EDA of transportation scores

```{r}
survey_ts |>
  pivot_longer(
    cols = c(rideshare_count, public_transit_count),
    names_to = "type",
    values_to = "count"
  ) |>
  ggplot(aes(x = count, fill = type)) +
  geom_density(alpha = 0.5)
```

```{r}
r2 <- summary(lm(rideshare_count ~ public_transit_count, data = survey_ts))$r.squared

ggplot(survey_ts, aes(x = rideshare_count, y = public_transit_count)) +
  geom_point() +
  geom_smooth(method = "lm") +
  annotate("text", x = Inf, y = Inf, hjust = 1.1, vjust = 1.5,
           label = paste0("R\u00b2 = ", round(r2, 2)))
```

```{r}
survey_ts |>
  mutate(highest_transport = highest_transport |>
           replace_na("NA") |>
           factor(levels = c("rideshare", "NA", "public_transit"))) |>
  ggplot(aes(x = highest_transport)) +
  geom_bar()
```
